---
title: "PECATI | Preliminary data monitoring, cleaning, and analysis. October 2023"
author: "Daniel Alcala"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(lubridate)
library(here)
library(cowplot)
library(gt)
library(gtsummary)
library(survival)
library(survminer)
library(epoxy)
library(crayon)

here("WP1_Preliminary-analysis_2023-10-24.Rmd")

cutoff_date <- as.Date("2023-10-24")
```

The data cutoff date for this analyses is: `r cutoff_date`

```{r, message = FALSE, warning = FALSE, echo = FALSE}
cutoff_date <- format(cutoff_date, "%Y_%m_%d")
```

# Patient Disposition
## Screening

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening <- read_excel(
    here(paste0("data/PECATI_ICF Screening_", cutoff_date, ".xlsx")),
    sheet = "ICF Screening",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    screening_date = "visit_date",
    eligibile = "is_the_patient_eligible_to_participate_in_the_study",
    sf_date = "non_eligible_date",
    sf_reason = "reason_not_eligible",
    eligibility_date_1 = "eligibility_date_v2",
    eligibility_date_2 = "eligibility_date_v3"
  ) %>% 
  mutate_at(
    vars(starts_with("eligibility_date")), dmy
  ) %>% 
  mutate(
    eligibility_date = if_else(
      eligibile == "Yes" & !is.na(eligibility_date_1), eligibility_date_1, eligibility_date_2),
    screening_date = dmy(screening_date),
    sf_date = dmy(sf_date),
    eligibility_date = ymd(eligibility_date)
  ) %>%
  select(
    patient, screening_date, eligibile, sf_date, sf_reason, eligibility_date
  )

screening_n <- length(data_00_screening$patient)
```

Listing of patients who have undergone the screening phase (n = `r screening_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing1_formatted_names <- c(
  "Patient",
  "Date of screening visit",
  "Eligible patient?",
  "Date of screening failure",
  "Reason for screening failure",
  "Date of confirmation of eligibility"
  )

data_00_screening_formatted <- data_00_screening %>% 
  set_names(listing1_formatted_names)

rmarkdown::paged_table(data_00_screening_formatted)
```

#### **Table 1. Patients included, excluded, and reasons for exclusion**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening %>% 
  tbl_summary(
    include = c(eligibile, sf_reason),
    missing = "no",
    label = list(
      eligibile ~ "Included",
      sf_reason ~ "Excluded"
    )
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Intention-to-treat population (ITT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_01_cycles <- read_excel(
  here(paste0("data/PECATI_Cycles_", cutoff_date, ".xlsx")),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names()

data_02_itt <- data_01_cycles %>% filter(
  event_num == 1
  ) %>% 
  select(patient)

data_temp_itt <- merge(
  data_02_itt,
  data_00_screening,
  by = "patient",
  all = FALSE
  ) %>% 
  select(patient, screening_date, eligibility_date)

itt_n <- length(data_temp_itt$patient)
```

Listing of patients who have started study treatment (n = `r itt_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing2_formatted_names <- c(
  "Patient",
  "Date of screening visit",
  "Date of confirmation of eligibility"
  )

data_temp_itt_formatted <- data_temp_itt %>% 
  set_names(listing2_formatted_names)

rmarkdown::paged_table(data_temp_itt_formatted)
```

## Study treatment discontinuation (EoT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_03_icf <- read_excel(
  here(paste0("data/PECATI_ICF Screening_", cutoff_date, ".xlsx")),
    sheet = "ICF",
    skip = 1) %>% 
  clean_names() %>% 
  rename(icf_date = "date_consent_obtained") %>% 
  select(patient, icf_date) %>% 
  mutate(
    icf_date = dmy(icf_date)
    )

data_03_icf <- data_03_icf %>% 
  group_by(patient) %>% 
  filter(icf_date == min(icf_date)) %>% 
  ungroup()

data_temp_icf <- merge(
  data_temp_itt,
  data_03_icf,
  by = "patient",
  all = FALSE  
  )

data_04_c1d1 <- read_excel(
  here(paste0("data/PECATI_Cycles_", cutoff_date, ".xlsx")),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    c1d1_date = "visit_date"
  ) %>% 
  mutate(
    c1d1_date = dmy(c1d1_date)
    ) %>% 
  filter(
    event_num == 1
  ) %>% 
  select(
    patient, c1d1_date
  )

data_temp_c1d1 <- merge(
  data_temp_icf,
  data_04_c1d1,
  by = "patient",
  all = FALSE  
)

data_05_eot <- read_excel(
  here(paste0("data/PECATI_FU30-EOT_", cutoff_date, ".xlsx")),
    sheet = "FU30-EOT",
    skip = 1
  ) %>% 
  clean_names() %>% 
  rename(
    eot_reason = "main_reason_for_treatment_discontinuation",
    last_pembrolizumab_date = "date_of_last_dose_of_pembrolizumab",
    last_lenvatinib_date = "date_of_last_dose_of_lenvatinib",
    radiological_pd = "disease_progression_confirmed_by_recist_v1_1",
    radiological_pd_date = "date_of_radiologial_progressive_disease",
    clinical_pd = "clinical_progression_in_accordance_to_investigator_criteria",
    clinical_pd_date = "date_of_clinical_progressive_disease"
  ) %>% 
  mutate(
    discontinuation_date = dmy(discontinuation_date),
    last_pembrolizumab_date = dmy(last_pembrolizumab_date),
    last_lenvatinib_date = dmy(last_lenvatinib_date),
    radiological_pd_date = dmy(radiological_pd_date),
    clinical_pd_date = dmy(clinical_pd_date)
    ) %>% 
  # Manual fixes
  mutate(
    eot_reason = if_else(
      patient == "0102-004" & eot_reason == "Other", "Second primary esophageal cancer", eot_reason)
  ) %>% 
  select(
    patient, discontinuation_date, eot_reason, last_pembrolizumab_date,
    last_lenvatinib_date, radiological_pd, radiological_pd_date, clinical_pd,
    clinical_pd_date
    )

data_temp_eot <- merge(
  data_temp_c1d1,
  data_05_eot,
  by = "patient",
  all = TRUE
  )

eot_n <- length(data_05_eot$patient)
```

Listing of patients who discontinued prematurely the study treatment (n = `r eot_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing3_formatted_names <- c(
  "Patient",
  "Date of end of treatment visit",
  "Reason for end of treatment",
  "Date of last dose of Pembrolizumab",
  "Date of last dose of Lenvantinib",
  "Radiological disease progression?",
  "Date of radiological disease progression",
  "Clinical disease progression?",
  "Date of clinical disease progression"
  )

data_05_eot_formatted <- data_05_eot %>% 
  set_names(listing3_formatted_names)

rmarkdown::paged_table(data_05_eot_formatted)
```


#### **Table 2. Patients who discontinued prematurely the study treatment:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_05_eot %>%
  tbl_summary(
    missing = "no",
    percent = "column",
    sort = list(everything() ~ "frequency"),
    label = list(eot_reason ~ "Reason for study treatment discontinuation"),
    include = c(eot_reason),
    missing_text = "Not reported"
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Study discontinuation (EoS)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_06_eos <- read_excel(
  here(paste0("data/PECATI_End of Study_", cutoff_date, ".xlsx")),
  sheet = "End of Study",
  skip = 1
  ) %>%
  clean_names() %>%
  rename(
    eos = "has_the_patient_continued_the_follow_up_until_the_eo_s_date",
    eos_date = "eo_s_date",
    eos_reason = "end_of_study_follow_up_main_reason",
    exitus_reason = "principal_cause_of_death"
  ) %>%
  select(patient, eos, eos_date, eos_reason, exitus_reason) %>%
  # Manual fixes
  mutate(
    eos_date = dmy(eos_date),
    exitus_reason = if_else(
      patient == "0102-003", "Disease progression", exitus_reason)
  )

data_temp_eos <- merge(
  data_temp_eot,
  data_06_eos,
  by = "patient",
  all = TRUE
)

eos_n <- length(data_06_eos$patient)
```

Listing of patients who discontinued the study (n = `r eos_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing4_formatted_names <- c(
  "Patient",
  "Did the patient complete the study?",
  "Date of end of study",
  "Resason for premature end of study",
  "Main reason for exitus"
  )

data_06_eos_formatted <- data_06_eos %>% 
  set_names(listing4_formatted_names)

rmarkdown::paged_table(data_06_eos_formatted)
```

## Swimmer plot

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_temp_eos <- data_temp_eos %>% 
  mutate(
    exitus = if_else(
      eos_reason == "Death", eos_date, as.Date(NA_real_)),
    progression = if_else(
      radiological_pd == "Yes", discontinuation_date, as.Date(NA_real_)),
  )

# Fill the NAs in the eos_date column with today's date
data_temp_eos$eos_date <- if_else(
  is.na(data_temp_eos$eos_date), as.Date(Sys.Date()), data_temp_eos$eos_date)

# Get the earliest date across columns for each patient and make it a date column
data_temp_eos$earliest_date <- apply(data_temp_eos[, sapply(data_temp_eos, is.Date)], 1, min, na.rm = TRUE)

# Get the latest date across columns for each patient and make it a date column
data_temp_eos$latest_date <- apply(data_temp_eos[, sapply(data_temp_eos, is.Date)], 1, max, na.rm = TRUE)

# Create a row for each patient for each day from the earliest to the latest date
data_07_swimmer_long <- data_temp_eos %>%
  mutate(
    earliest_date = ymd(earliest_date),
    latest_date = ymd(latest_date)
  ) %>% 
  rowwise() %>%
  mutate(study_day = list(seq.Date(from = earliest_date, to = latest_date, by = "day"))) %>%
  unnest(study_day)

# Count the number of rows for each patient
data_07_swimmer_long <- data_07_swimmer_long %>%
  group_by(patient) %>%
  mutate(event_num = row_number())

data_07_swimmer_long <- data_07_swimmer_long %>% 
  mutate(
    month = event_num * 0.0328767
  )

data_07_swimmer_long$icf_this_day <- ifelse(
  data_07_swimmer_long$icf_date == data_07_swimmer_long$study_day, data_07_swimmer_long$month, NA
  )

data_07_swimmer_long$c1d1_this_day <- ifelse(
  data_07_swimmer_long$c1d1_date == data_07_swimmer_long$study_day, data_07_swimmer_long$month, NA
  )

data_07_swimmer_long$pd_this_day <- ifelse(
  data_07_swimmer_long$progression == data_07_swimmer_long$study_day, data_07_swimmer_long$month, NA
  )

data_07_swimmer_long$exitus_this_day <- ifelse(
  data_07_swimmer_long$exitus == data_07_swimmer_long$study_day, data_07_swimmer_long$month, NA
  )

data_07_swimmer_long <- data_07_swimmer_long %>% 
  group_by(patient) %>% 
  mutate(max_day = max(event_num)) %>% 
  ungroup() %>% 
  mutate(patient = fct_reorder(factor(patient), max_day))
```

#### **Figure 1. Patient trajectories along the clinical trial from informed consent to study termination**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
cols <- c(
  "Study treatment start date" = "#74AAFF",
  "Disease progression" = "#FFD966",
  "Exitus" = "#040404"
)

shape_override <- c(21, 21, 21)
stroke_override <- c(1, 1, 1)
size_override <-  c(2, 2, 2)

swimmer <- data_07_swimmer_long %>%
  ggplot(
    aes(y = patient, group = patient)) +
  theme_bw() + 
  theme(
    panel.grid.minor.x = element_blank(),
    panel.border = element_blank()
  ) +
  # 1 data layer: line plot showing number of TAs
  geom_line(aes(x = month), linewidth = 2.5, color = "#9A99A0", alpha = 0.25) +
  # 2 data layer: dot plot showing when the ICF was signed
  geom_point(aes(x = icf_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#74AAFF",
             fill = "#74AAFF",
             alpha = 0.65) +
  # 3 data layer: dot plot showing the study treatment start date
  geom_point(aes(x = c1d1_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 1,
             color = "#74AAFF") +
  # 4 data layer: dot plot showing whether disease progression
  geom_point(aes(x = pd_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#040404",
             fill = "#FFD966",
             alpha = 0.95) +
  # 5 data layer: dot plot showing TAs with an OR = SD
  geom_point(aes(x = exitus_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#040404",
             fill = "#040404") +
  # Final aesthetics adjustments
  scale_color_manual(values = cols,
                     name = "XXXX") +
  scale_x_continuous(breaks = seq(0, 36, 5)) +
  guides(color = guide_legend(
  override.aes = list(
    shape = shape_override,
    stroke = stroke_override,
    size = size_override
  )
)) +
  labs(
    x = "Time since ICF sign (months)",
    y = "Patient"
  ) +
  theme(panel.grid.major.y = element_blank())

ggsave(
  here("output/PECATI_swimmer_2023-10-24.png"),
  swimmer,
  width = 2048,
  height = 4096,
  units = "px",
  dpi = 300
  )
  
swimmer
```

# Tumour Assessment

This section presents the cleaning of tumour assessment data from the PECATI study. The ultimate goal is to obtain the number of patients who have experienced radiological disease progression. However, the main purpose of this preliminary analysis of radiological progressions is to detect anomalies or discrepancies in the tumour assessment data collected.

Data from tumour assessments need to be pulled from two different sources. On the one hand, we have the baseline tumour assessment included in the Screening sheet. On the other hand, we have the tumour assessment sheet itself, which collects all post-baseline assessments.

In addition, some patients might not have measurable lesions at baseline. Therefore, we need to collect data for both types of lesions. 

## Baseline tumor assessment
### Measurable disease

First, we extract data from patients with measurable lesions at baseline:


```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_08a_baseline_target <- read_excel(
    here(paste0("data/PECATI_Tumor Assessment_", cutoff_date, ".xlsx")),
    sheet = "TARGET",
    skip = 1) %>% 
    clean_names() %>% 
    # The variable "event_num" encodes the tumour assessment number in the post-baseline assessments.
    # Create this variable and assign a 0 to mark this as the baseline assessment.
    mutate(
      event_num = 0
    ) %>%
    group_by(patient, event_num) %>%
    # We collect the length of the longest diameter of each reported target (i.e., measurable) lesion individually.
    # Calculate the sum of longest diameter of all lesions.
    mutate(
      sum_of_lesions = sum(longest_diameter_short_axis_mm)
    ) %>%
    # Since we now have the sum of longest diameters repeated as many times as there are measurable lesions, we can keep either any row
    group_by(patient) %>%
    filter(row_number() == 1) %>%
    mutate(
      evaluation_date = dmy(evaluation_date)
    ) %>% 
    # Create a dummy variable to code that the patient had measurable disease at baseline.
    mutate(
      baseline_target = 0
    ) %>% 
    select(
        patient, event_num, baseline_target, evaluation_date, longest_diameter_short_axis_mm, sum_of_lesions
        )

data_08a_baseline_target <- merge(
  data_02_itt,
  data_08a_baseline_target,
  by = "patient",
  all = FALSE
)

baseline_target_n <- length(data_08a_baseline_target$patient)
```

Listing of patients with measurable disease at baseline phase (n = `r baseline_target_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing5_formatted_names <- c(
  "Patient",
  "Date of baseline tumor assessment",
  "Longest diameter (short axis) in mm",
  "Sum of longest diameters (SLDs)"
  )

data_08a_baseline_target_formatted <- data_08a_baseline_target %>% 
  select(patient, evaluation_date, longest_diameter_short_axis_mm, sum_of_lesions) %>% 
  set_names(listing5_formatted_names)
  
rmarkdown::paged_table(data_08a_baseline_target_formatted)
```

### Non-measurable disease

Then, we extract data from non-measurable lesions at baseline for all patients:

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_08b_baseline_nontarget <- read_excel(
    here(paste0("data/PECATI_Tumor Assessment_", cutoff_date, ".xlsx")),
    sheet = "NON_TARGET",
    skip = 1) %>% 
    clean_names() %>% 
    # The variable "event_num" encodes the tumour assessment number in the post-baseline assessments.
    # Create this variable and assign a 0 to mark this as the baseline assessment.
    mutate(
      event_num = 0
    ) %>%
    # Since we only care about the number of patients with non-measurable disease at baseline, we can keep either any row
    group_by(patient) %>%
    filter(row_number() == 1) %>%
    mutate(
      evaluation_date = dmy(evaluation_date)
    ) %>% 
    select(
      patient, event_num, evaluation_date
    )

data_08b_baseline_nontarget <- merge(
  data_02_itt,
  data_08b_baseline_nontarget,
  by = "patient",
  all = FALSE
)

baseline_nontarget_n <- length(data_08b_baseline_nontarget$patient)
```

Listing of patients with measurable disease at baseline phase (n = `r baseline_nontarget_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing6_formatted_names <- c(
  "Patient",
  "Date of baseline tumor assessment"
  )

data_08b_baseline_nontarget_formatted <- data_08b_baseline_nontarget %>% 
  select(patient, evaluation_date) %>% 
  set_names(listing6_formatted_names)

rmarkdown::paged_table(data_08b_baseline_nontarget_formatted)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_08_baseline_all <- full_join(
  data_08a_baseline_target,
  data_08b_baseline_nontarget,
  by = c("patient", "event_num")
  ) %>%
  # If a patient had measurable lesions at baseline, we want to keep that data.
  # Only if there were no measurable lesions, we want to keep the tumour assessment date when the non-measurable lesions were detected.
  mutate(
    evaluation_date = if_else(
      is.na(evaluation_date.x), evaluation_date.y, evaluation_date.x
    )
  ) %>%
  # The following mutate() function creates a "baseline_nontarget" variable that codes 0 if the patient had no measurable lesions at baseline.
  mutate(
    baseline_nontarget = if_else(is.na(baseline_target) == 0, 1, 0),
    baseline_nontarget = na_if(baseline_nontarget, 1)
  ) %>% 
  select(
    patient, event_num, evaluation_date, baseline_target, sum_of_lesions, baseline_nontarget
  )

# rmarkdown::paged_table(data_08_baseline_all)
```

### Patients without baseline tumour assessment

```{r, message = FALSE, warning = FALSE, echo = FALSE}
qa_no_baseline_ta <- anti_join(
  data_02_itt,
  data_08_baseline_all,
  by = "patient"
)

no_baseline_ta_n <- length(qa_no_baseline_ta$patient)
```

There are `r no_baseline_ta_n` patients in the ITT population who do not have a baseline tumor assessment:

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing7_formatted_names <- c(
  "Patient"
  )

qa_no_baseline_ta_formatted <- qa_no_baseline_ta %>% 
  select(patient) %>% 
  set_names(listing7_formatted_names)

rmarkdown::paged_table(qa_no_baseline_ta)
```

### Quality assurance of baseline tumor burden

A histogram with the sum of the longest diameters of each patient is presented below with the intention of detecting possible strange or unlikely patterns:

```{r, message = FALSE, warning = FALSE, echo = FALSE}
ggplot(
  data_08_baseline_all, 
  aes(x = sum_of_lesions)) +
  geom_histogram(
    fill = "#F3656C", 
    alpha = 0.5, 
    color = "#F3656C",
    bins = 54) +
  theme(
    panel.background = element_rect(fill = "#FFFFFF", colour = "#FFFFFF"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background  = element_rect(fill = "#FFFFFF", colour = "#FFFFFF")
  ) +
  labs(
    x = "Sum of longest diameters at baseline (mm)",
    y = "Number of patients"
  )
```

# Preliminary Efficacy Analysis

# Preliminary Safety Analysis

