---
title: "PECATI | Preliminary data monitoring, cleaning, and analysis. October 2023"
author: "Daniel Alcala"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(lubridate)
library(here)
library(cowplot)
library(gt)
library(gtsummary)
library(survival)
library(survminer)
library(epoxy)
library(crayon)

here("WP1_Preliminary-analysis_2023-10-24.Rmd")

cutoff_date <- as.Date("2023-10-24")
cutoff_date <- format(cutoff_date, "%Y_%m_%d")
```

# Patient Disposition
## Screening

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening <- read_excel(
    here(paste0("data/PECATI_ICF Screening_", cutoff_date, ".xlsx")),
    sheet = "ICF Screening",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    screening_date = "visit_date",
    eligibile = "is_the_patient_eligible_to_participate_in_the_study",
    sf_date = "non_eligible_date",
    sf_reason = "reason_not_eligible",
    eligibility_date_1 = "eligibility_date_v2",
    eligibility_date_2 = "eligibility_date_v3"
  ) %>% 
  mutate_at(
    vars(starts_with("eligibility_date")), dmy
  ) %>% 
  mutate(
    eligibility_date = if_else(
      eligibile == "Yes" & !is.na(eligibility_date_1), eligibility_date_1, eligibility_date_2),
    screening_date = dmy(screening_date),
    sf_date = dmy(sf_date),
    eligibility_date = ymd(eligibility_date)
  ) %>%
  select(
    patient, screening_date, eligibile, sf_date, sf_reason, eligibility_date
  )

screening_n <- length(data_00_screening$patient)
```

Listing of patients who have undergone the screening phase (n = `r screening_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing1_formatted_names <- c(
  "Patient",
  "Date of screening visit",
  "Eligible patient?",
  "Date of screening failure",
  "Reason for screening failure",
  "Date of confirmation of eligibility"
  )

data_00_screening_formatted <- data_00_screening %>% 
  set_names(listing1_formatted_names)

rmarkdown::paged_table(data_00_screening_formatted)
```

#### **Table 1. Patients included, excluded, and reasons for exclusion**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening %>% 
  tbl_summary(
    include = c(eligibile, sf_reason),
    missing = "no",
    label = list(
      eligibile ~ "Included",
      sf_reason ~ "Excluded"
    )
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Intention-to-treat population (ITT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_01_cycles <- read_excel(
  here(paste0("data/PECATI_Cycles_", cutoff_date, ".xlsx")),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names()

data_02_itt <- data_01_cycles %>% filter(
  event_num == 1
  ) %>% 
  select(patient)

data_temp_itt <- merge(
  data_02_itt,
  data_00_screening,
  by = "patient",
  all = FALSE
  ) %>% 
  select(patient, screening_date, eligibility_date)

itt_n <- length(data_temp_itt$patient)
```

Listing of patients who have started study treatment (n = `r itt_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing2_formatted_names <- c(
  "Patient",
  "Date of screening visit",
  "Date of confirmation of eligibility"
  )

data_temp_itt_formatted <- data_temp_itt %>% 
  set_names(listing2_formatted_names)

rmarkdown::paged_table(data_temp_itt_formatted)
```

## Study treatment discontinuation (EoT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_03_icf <- read_excel(
  here(paste0("data/PECATI_ICF Screening_", cutoff_date, ".xlsx")),
    sheet = "ICF",
    skip = 1) %>% 
  clean_names() %>% 
  rename(icf_date = "date_consent_obtained") %>% 
  select(patient, icf_date) %>% 
  mutate(
    icf_date = dmy(icf_date)
    )

data_03_icf <- data_03_icf %>% 
  group_by(patient) %>% 
  filter(icf_date == min(icf_date)) %>% 
  ungroup()

data_temp_icf <- merge(
  data_temp_itt,
  data_03_icf,
  by = "patient",
  all = FALSE  
  )

data_04_c1d1 <- read_excel(
  here(paste0("data/PECATI_Cycles_", cutoff_date, ".xlsx")),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    c1d1_date = "visit_date"
  ) %>% 
  mutate(
    c1d1_date = dmy(c1d1_date)
    ) %>% 
  filter(
    event_num == 1
  ) %>% 
  select(
    patient, c1d1_date
  )

data_temp_c1d1 <- merge(
  data_temp_icf,
  data_04_c1d1,
  by = "patient",
  all = FALSE  
)

data_05_eot <- read_excel(
  here(paste0("data/PECATI_FU30-EOT_", cutoff_date, ".xlsx")),
    sheet = "FU30-EOT",
    skip = 1
  ) %>% 
  clean_names() %>% 
  rename(
    eot_reason = "main_reason_for_treatment_discontinuation",
    last_pembrolizumab_date = "date_of_last_dose_of_pembrolizumab",
    last_lenvatinib_date = "date_of_last_dose_of_lenvatinib",
    radiological_pd = "disease_progression_confirmed_by_recist_v1_1",
    radiological_pd_date = "date_of_radiologial_progressive_disease",
    clinical_pd = "clinical_progression_in_accordance_to_investigator_criteria",
    clinical_pd_date = "date_of_clinical_progressive_disease"
  ) %>% 
  mutate(
    discontinuation_date = dmy(discontinuation_date),
    last_pembrolizumab_date = dmy(last_pembrolizumab_date),
    last_lenvatinib_date = dmy(last_lenvatinib_date),
    radiological_pd_date = dmy(radiological_pd_date),
    clinical_pd_date = dmy(clinical_pd_date)
    ) %>% 
  # Manual fixes
  mutate(
    eot_reason = if_else(
      patient == "0102-004" & eot_reason == "Other", "Second primary esophageal cancer", eot_reason)
  ) %>% 
  select(
    patient, discontinuation_date, eot_reason, last_pembrolizumab_date,
    last_lenvatinib_date, radiological_pd, radiological_pd_date, clinical_pd,
    clinical_pd_date
    )

data_temp_eot <- merge(
  data_temp_c1d1,
  data_05_eot,
  by = "patient",
  all = TRUE
  )

eot_n <- length(data_05_eot$patient)
```

Listing of patients who discontinued prematurely the study treatment (n = `r eot_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing3_formatted_names <- c(
  "Patient",
  "Date of end of treatment visit",
  "Reason for end of treatment",
  "Date of last dose of Pembrolizumab",
  "Date of last dose of Lenvantinib",
  "Radiological disease progression?",
  "Date of radiological disease progression",
  "Clinical disease progression?",
  "Date of clinical disease progression"
  )

data_05_eot_formatted <- data_05_eot %>% 
  set_names(listing3_formatted_names)

rmarkdown::paged_table(data_05_eot_formatted)
```


#### **Table 2. Patients who discontinued prematurely the study treatment:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_05_eot %>%
  tbl_summary(
    missing = "no",
    percent = "column",
    sort = list(everything() ~ "frequency"),
    label = list(eot_reason ~ "Reason for study treatment discontinuation"),
    include = c(eot_reason),
    missing_text = "Not reported"
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Study discontinuation (EoS)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_06_eos <- read_excel(
  here(paste0("data/PECATI_End of Study_", cutoff_date, ".xlsx")),
  sheet = "End of Study",
  skip = 1
  ) %>%
  clean_names() %>%
  rename(
    eos = "has_the_patient_continued_the_follow_up_until_the_eo_s_date",
    eos_date = "eo_s_date",
    eos_reason = "end_of_study_follow_up_main_reason",
    exitus_reason = "principal_cause_of_death"
  ) %>%
  select(patient, eos, eos_date, eos_reason, exitus_reason) %>%
  # Manual fixes
  mutate(
    eos_date = dmy(eos_date),
    exitus_reason = if_else(
      patient == "0102-003", "Disease progression", exitus_reason)
  )

data_temp_eos <- merge(
  data_temp_eot,
  data_06_eos,
  by = "patient",
  all = TRUE
)

eos_n <- length(data_06_eos$patient)
```

Listing of patients who discontinued the study (n = `r eos_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
listing4_formatted_names <- c(
  "Patient",
  "Did the patient complete the study?",
  "Date of end of study",
  "Resason for premature end of study",
  "Main reason for exitus"
  )

data_06_eos_formatted <- data_06_eos %>% 
  set_names(listing4_formatted_names)

rmarkdown::paged_table(data_06_eos_formatted)
```

## Swimmer plot

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_temp_eos <- data_temp_eos %>% 
  mutate(
    exitus = if_else(
      eos_reason == "Death", eos_date, as.Date(NA_real_)),
    progression = if_else(
      radiological_pd == "Yes", discontinuation_date, as.Date(NA_real_)),
  )

# Fill the NAs in the eos_date column with today's date
data_temp_eos$eos_date <- if_else(
  is.na(data_temp_eos$eos_date), as.Date(Sys.Date()), data_temp_eos$eos_date)

# Get the earliest date across columns for each patient and make it a date column
data_temp_eos$earliest_date <- apply(data_temp_eos[, sapply(data_temp_eos, is.Date)], 1, min, na.rm = TRUE)

# Get the latest date across columns for each patient and make it a date column
data_temp_eos$latest_date <- apply(data_temp_eos[, sapply(data_temp_eos, is.Date)], 1, max, na.rm = TRUE)

# Create a row for each patient for each day from the earliest to the latest date
data_08_swimmer_long <- data_temp_eos %>%
  mutate(
    earliest_date = ymd(earliest_date),
    latest_date = ymd(latest_date)
  ) %>% 
  rowwise() %>%
  mutate(study_day = list(seq.Date(from = earliest_date, to = latest_date, by = "day"))) %>%
  unnest(study_day)

# Count the number of rows for each patient
data_08_swimmer_long <- data_08_swimmer_long %>%
  group_by(patient) %>%
  mutate(event_num = row_number())

data_08_swimmer_long <- data_08_swimmer_long %>% 
  mutate(
    month = event_num * 0.0328767
  )

data_08_swimmer_long$icf_this_day <- ifelse(
  data_08_swimmer_long$icf_date == data_08_swimmer_long$study_day, data_08_swimmer_long$month, NA
  )

data_08_swimmer_long$c1d1_this_day <- ifelse(
  data_08_swimmer_long$c1d1_date == data_08_swimmer_long$study_day, data_08_swimmer_long$month, NA
  )

data_08_swimmer_long$pd_this_day <- ifelse(
  data_08_swimmer_long$progression == data_08_swimmer_long$study_day, data_08_swimmer_long$month, NA
  )

data_08_swimmer_long$exitus_this_day <- ifelse(
  data_08_swimmer_long$exitus == data_08_swimmer_long$study_day, data_08_swimmer_long$month, NA
  )

data_08_swimmer_long <- data_08_swimmer_long %>% 
  group_by(patient) %>% 
  mutate(max_day = max(event_num)) %>% 
  ungroup() %>% 
  mutate(patient = fct_reorder(factor(patient), max_day))
```

#### **Figure 1. Patient trajectories along the clinical trial from informed consent to study termination**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
cols <- c(
  "Study treatment start date" = "#74AAFF",
  "Disease progression" = "#FFD966",
  "Exitus" = "#040404"
)

shape_override <- c(21, 21, 21)
stroke_override <- c(1, 1, 1)
size_override <-  c(2, 2, 2)

swimmer <- data_08_swimmer_long %>%
  ggplot(
    aes(y = patient, group = patient)) +
  theme_bw() + 
  theme(
    panel.grid.minor.x = element_blank(),
    panel.border = element_blank()
  ) +
  # 1 data layer: line plot showing number of TAs
  geom_line(aes(x = month), linewidth = 2.5, color = "#9A99A0", alpha = 0.25) +
  # 2 data layer: dot plot showing when the ICF was signed
  geom_point(aes(x = icf_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#74AAFF",
             fill = "#74AAFF",
             alpha = 0.65) +
  # 3 data layer: dot plot showing the study treatment start date
  geom_point(aes(x = c1d1_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 1,
             color = "#74AAFF") +
  # 4 data layer: dot plot showing whether disease progression
  geom_point(aes(x = pd_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#040404",
             fill = "#FFD966",
             alpha = 0.95) +
  # 5 data layer: dot plot showing TAs with an OR = SD
  geom_point(aes(x = exitus_this_day),
             size = 2.5,
             stroke = 0.5,
             shape = 21,
             color = "#040404",
             fill = "#040404") +
  # Final aesthetics adjustments
  scale_color_manual(values = cols,
                     name = "XXXX") +
  scale_x_continuous(breaks = seq(0, 36, 5)) +
  guides(color = guide_legend(
  override.aes = list(
    shape = shape_override,
    stroke = stroke_override,
    size = size_override
  )
)) +
  labs(
    x = "Time since ICF sign (months)",
    y = "Patient"
  ) +
  theme(panel.grid.major.y = element_blank())

ggsave(
  here("output/PECATI_swimmer_2023-10-24.png"),
  swimmer,
  width = 2048,
  height = 4096,
  units = "px",
  dpi = 300
  )
  
swimmer
```

