---
title: "PECATI | Preliminary data monitoring, cleaning, and analysis. October 2023"
author: "Daniel Alcala"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(lubridate)
library(here)
library(cowplot)
library(gt)
library(gtsummary)
library(survival)
library(survminer)
library(epoxy)
library(crayon)

here("WP1_Preliminary-analysis_2023-10-24.Rmd")
```

# Patient Disposition
## Screening

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening <- read_excel(
  here("data/PECATI_ICF Screening_2023_10_24.xlsx"),
    sheet = "ICF Screening",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    screening_date = "visit_date",
    eligibile = "is_the_patient_eligible_to_participate_in_the_study",
    sf_date = "non_eligible_date",
    sf_reason = "reason_not_eligible",
    eligibility_date_1 = "eligibility_date_v2",
    eligibility_date_2 = "eligibility_date_v3"
  ) %>% 
  mutate_at(
    vars(starts_with("eligibility_date")), dmy
  ) %>% 
  mutate(
    eligibility_date = if_else(
      eligibile == "Yes" & !is.na(eligibility_date_1), eligibility_date_1, eligibility_date_2)
  ) %>%
  select(
    patient, screening_date, eligibile, sf_date, sf_reason, eligibility_date
  )

screening_n <- length(data_00_screening$patient)
```

Listing of patients who have undergone the screening phase (n = `r screening_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
rmarkdown::paged_table(data_00_screening)
```

#### **Table 1. Patients included, excluded, and reasons for exclusion**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_00_screening %>% 
  tbl_summary(
    include = c(eligibile, sf_reason),
    missing = "no",
    label = list(
      eligibile ~ "Included",
      sf_reason ~ "Excluded"
    )
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Intention-to-treat population (ITT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_01_cycles <- read_excel(
  here("data/PECATI_Cycles_2023_10_24.xlsx"),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names()

data_02_itt <- data_01_cycles %>% filter(
  event_num == 1
  ) %>% 
  select(patient)

data_temp_itt <- merge(
  data_02_itt,
  data_00_screening,
  by = "patient",
  all = FALSE
  ) %>% 
  select(patient, screening_date, eligibility_date)

itt_n <- length(data_temp_itt$patient) 
```

Listing of patients who have started study treatment (n = `r itt_n`):

```{r, message = FALSE, warning = FALSE, echo = FALSE}
rmarkdown::paged_table(data_temp_itt)
```

## Study treatment discontinuations (EoT)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_03_icf <- read_excel(
  here("data/PECATI_ICF Screening_2023_10_24.xlsx"),
    sheet = "ICF",
    skip = 1) %>% 
  clean_names() %>% 
  rename(icf_date = "date_consent_obtained") %>% 
  select(patient, icf_date) %>% 
  mutate(
    icf_date = dmy(icf_date)
    )

data_03_icf <- data_03_icf %>% 
  group_by(patient) %>% 
  filter(icf_date == min(icf_date)) %>% 
  ungroup()

data_temp_icf <- merge(
  data_temp_itt,
  data_03_icf,
  by = "patient",
  all = FALSE  
  )

data_04_c1d1 <- read_excel(
  here("data/PECATI_Cycles_2023_10_24.xlsx"),
    sheet = "Cycles",
    skip = 1) %>% 
  clean_names() %>% 
  rename(
    c1d1_date = "visit_date"
  ) %>% 
  mutate(
    c1d1_date = dmy(c1d1_date)
    ) %>% 
  filter(
    event_num == 1
  ) %>% 
  select(
    patient, c1d1_date
  )

data_temp_c1d1 <- merge(
  data_temp_icf,
  data_04_c1d1,
  by = "patient",
  all = FALSE  
)

data_05_eot <- read_excel(
  here("data/PECATI_FU30-EOT_2023_10_24.xlsx"), 
    sheet = "FU30-EOT",
    skip = 1
  ) %>% 
  clean_names() %>% 
  rename(
    eot_reason = "main_reason_for_treatment_discontinuation",
    last_pembrolizumab_date = "date_of_last_dose_of_pembrolizumab",
    last_lenvatinib_date = "date_of_last_dose_of_lenvatinib",
    radiological_pd = "disease_progression_confirmed_by_recist_v1_1",
    radiological_pd_date = "date_of_radiologial_progressive_disease",
    clinical_pd = "clinical_progression_in_accordance_to_investigator_criteria",
    clinical_pd_date = "date_of_clinical_progressive_disease"
  ) %>% 
  mutate(
    discontinuation_date = dmy(discontinuation_date),
    last_pembrolizumab_date = dmy(last_pembrolizumab_date),
    last_lenvatinib_date = dmy(last_lenvatinib_date),
    radiological_pd_date = dmy(radiological_pd_date),
    clinical_pd_date = dmy(clinical_pd_date)
    ) %>% 
  # Manual fixes
  mutate(
    eot_reason = if_else(
      patient == "0102-004" & eot_reason == "Other", "Second primary esophageal cancer", eot_reason)
  ) %>% 
  select(
    patient, discontinuation_date, eot_reason, last_pembrolizumab_date,
    last_lenvatinib_date, radiological_pd, radiological_pd_date, clinical_pd,
    clinical_pd_date
    )
```

#### **Table 2. Patients who discontinued prematurely the study treatment:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_05_eot %>%
  tbl_summary(
    missing = "no",
    percent = "column",
    sort = list(everything() ~ "frequency"),
    label = list(eot_reason ~ "Reason for study treatment discontinuation"),
    include = c(eot_reason),
    missing_text = "Not reported"
  ) %>%
  modify_header(label = " ") %>%
  bold_labels()
```

## Study discontinuation (EoS)

## Swimmer plot





