---
title: "PECATI | Data monitoring, cleaning, and analysis for the presentation at ESMO 2024. August 2024"
author: "Daniel Alcala"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(lubridate)
library(here)
library(cowplot)
library(gt)
library(gtsummary)
library(gtreg)
library(survival)
library(survminer)
library(epoxy)
library(crayon)
library(grid)
library(workflowr)
library(broom)
library(labelled)

here("WP1_preliminary-analysis.Rmd")

project_id <- "PECATI"
cutoff_date <- as.Date("2024-07-17")
```

The data cutoff date for these analyses is: `r cutoff_date`

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Reformat the data cutoff date for reading of files
cutoff_date_formatted <- format(cutoff_date, "%Y_%m_%d")
```

# Efficacy Analysis

## Progression-free survival (PFS)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
pfs5mo <- read_excel(
  here(paste0("data/", "MedOPP341_PECATI_Draft_Listings_", cutoff_date, ".xlsx")),
  sheet = "LIST18"
) %>%
  clean_names() %>% 
  rename(
    patient = subject_id,
    pfs_event = pfs_at_5_months,
    pfs = time_of_pfs_months
  ) %>% 
  ### Manual fix ###-----------------------------------------------------------
  mutate(
    ### Patient 0206-002 had a PFS event identified at the second assessment,
    ### which was at 7.49 months (2023-12-28), due to the identification of a 
    ### new lesion. However, the new lesion was not confirmed at the next
    ### assessment, and the patient continued on treatment.
    pfs_event = if_else(
      patient == "0206-002", "No", pfs_event
    ),
    pfs = if_else(
      patient == "0206-002", 7.49, pfs
    ),
    pfs_event = if_else(
      pfs_event == "No", 0, 1
    )
  )
```

#### **Table 1. 5-month PFS:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
tbl_survfit(
  survfit(Surv(pfs, pfs_event) ~ 1, data = pfs5mo, conf.type = "log-log"),
  times = c(5),
  label_header = "**{time}-Month PFS**"
) %>%
  modify_header(label = " ")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
pfs <- read_excel(
  here(paste0("data/", "MedOPP341_PECATI_Draft_Listings_", cutoff_date, ".xlsx")),
  sheet = "LIST18"
) %>%
  clean_names() %>% 
  rename(
    patient = subject_id,
    pfs_event = progression_free_survival_pfs,
    pfs = time_of_pfs_months
  ) %>% 
  mutate(
    pfs_event = if_else(
      patient == "0206-002", "No", pfs_event
    ),
    pfs = if_else(
      patient == "0206-002", 7.49, pfs
    ),
    pfs_event = if_else(
      pfs_event == "No", 0, 1
    )
  )
```

#### **Table 2. 6-, 12-, and 24-month Progression-Free Survival (PFS):**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
tbl_survfit(
  survfit(Surv(pfs, pfs_event) ~ 1, data = pfs, conf.type = "log-log"),
  times = c(6, 12, 24),
  label_header = "**{time}-Month PFS**"
) %>%
  modify_header(label = " ")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
customize_labels <- function(p, font.title = NULL,
                             font.subtitle = NULL, font.caption = NULL,
                             font.x = NULL, font.y = NULL, font.xtickslab = NULL, font.ytickslab = NULL) {
  original.p <- p
  if (is.ggplot(original.p)) {
    list.plots <- list(original.p)
  } else if (is.list(original.p)) {
    list.plots <- original.p
  } else {
    stop("Can't handle an object of class ", class(original.p))
  }
  .set_font <- function(font) {
    font <- ggpubr:::.parse_font(font)
    ggtext::element_markdown(size = font$size, face = font$face, colour = font$color)
  }
  for (i in 1:length(list.plots)) {
    p <- list.plots[[i]]
    if (is.ggplot(p)) {
      if (!is.null(font.title)) p <- p + theme(plot.title = .set_font(font.title))
      if (!is.null(font.subtitle)) p <- p + theme(plot.subtitle = .set_font(font.subtitle))
      if (!is.null(font.caption)) p <- p + theme(plot.caption = .set_font(font.caption))
      if (!is.null(font.x)) p <- p + theme(axis.title.x = .set_font(font.x))
      if (!is.null(font.y)) p <- p + theme(axis.title.y = .set_font(font.y))
      if (!is.null(font.xtickslab)) p <- p + theme(axis.text.x = .set_font(font.xtickslab))
      if (!is.null(font.ytickslab)) p <- p + theme(axis.text.y = .set_font(font.ytickslab))
      list.plots[[i]] <- p
    }
  }
  if (is.ggplot(original.p)) {
    list.plots[[1]]
  } else {
    list.plots
  }
}
```

#### **Figure 1. Progression-free survival (PFS) in the intention-to-treat population**

```{r pfs-kaplan-meier-plot, message = FALSE, warning = FALSE, echo = FALSE}
### Plot ----------------------------------------------------------------------
primary_endpoint <- ggsurvplot(
  fit = survfit(Surv(pfs, pfs_event) ~ 1, data = pfs),
  xlab = "Months since treatment initiation",
  ylab = "Progression-Free Survival",
  conf.int = FALSE,
  xlim = c(0, 24.9),
  ylim = c(0, 1.009),
  censor.size = 3.5,
  size = 1,
  break.time.by = 3,
  axes.offset = FALSE,
  surv.scale = "percent",
  break.y.by = 0.20,
  risk.table = TRUE,
  fontsize = 2.5,
  ggtheme = theme_classic(),
  palette = c("#0A3158"),
  legend.labs = c(" "),
  legend.title = " "
)

### Modify theme --------------------------------------------------------------
primary_endpoint <- customize_labels(
  primary_endpoint,
  font.title     = c(8, "bold", "#0a0908"),
  font.caption   = c(8, "plain", "#0a0908"),
  font.x         = c(8, "bold", "#0a0908"),
  font.y         = c(8, "bold", "#0a0908"),
  font.xtickslab = c(8, "plain", "#0a0908"),
  font.ytickslab = c(8, "plain", "#0a0908")
)

grid.draw.ggsurvplot <- function(x) {
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

### Add legend ----------------------------------------------------------------
caption <- paste(strwrap(
  "5-month PFS is 87.8% (95% CI 78.3-98.4). Events: 5/43 (11.6%).", 90
  ), collapse = "\n")

primary_endpoint$plot <- primary_endpoint$plot + annotate(
  "text",
  x = 0.5, y = 0.10,
  label = caption,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

### Save plot -----------------------------------------------------------------
ggsave(
  paste("../output/", project_id, "_5mo_PFS_global_", as.Date(Sys.Date()), ".png", sep = ""),
  primary_endpoint,
  width = 20,
  height = 10,
  units = "cm",
  dpi = 300
)

### Show plot -----------------------------------------------------------------
primary_endpoint
```

#### **Table 3. PFS by PD-L1 status:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
var_label(pfs$pd_l1) <- "PD-L1 expression"

tbl_survfit(
  survfit(Surv(pfs, pfs_event) ~ pd_l1, data = pfs, conf.type = "log-log"),
  times = c(5, 12),
  label_header = "**{time}-Month PFS**"
) %>%
  modify_header(label = " ")
```

#### **Figure 2. Progression-Free Survival (PFS) by PD-L1 status in the intention-to-treat population**

```{r pfs-pdl1-kaplan-meier-plot, message = FALSE, warning = FALSE, echo = FALSE}
### Plot ----------------------------------------------------------------------
pfs_pdl1 <- ggsurvplot(
  fit = survfit(Surv(pfs, pfs_event) ~ pd_l1, data = pfs),
  xlab = "Months since treatment initiation",
  ylab = "Progression-Free Survival",
  conf.int = FALSE,
  pval = TRUE,
  pval.size = 3,
  pval.coord = c(0.5, 0.28),
  xlim = c(0, 24.9),
  ylim = c(0, 1.009),
  censor.size = 3.5,
  size = 1,
  break.time.by = 3,
  axes.offset = FALSE,
  surv.scale = "percent",
  break.y.by = 0.20,
  risk.table = TRUE,
  fontsize = 2.5,
  ggtheme = theme_classic(),
  palette = c("#4F5D67", "#BB322F"),
  legend.labs = c("PD-L1[-]", "PD-L1[+]"),
  legend.title = " "
)

### Modify theme --------------------------------------------------------------
pfs_pdl1 <- customize_labels(
  pfs_pdl1,
  font.title     = c(8, "bold", "#0a0908"),
  font.caption   = c(8, "plain", "#0a0908"),
  font.x         = c(8, "bold", "#0a0908"),
  font.y         = c(8, "bold", "#0a0908"),
  font.xtickslab = c(8, "plain", "#0a0908"),
  font.ytickslab = c(8, "plain", "#0a0908")
)

grid.draw.ggsurvplot <- function(x) {
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

### Add legend ----------------------------------------------------------------
caption2 <- paste(strwrap(
  "5-month PFS is 93.8% (95% CI 82.6-100) in PD-L1 negative patients. Events: 1/17 (5.9%).", 90
  ), collapse = "\n")

caption3 <- paste(strwrap(
  "5-month PFS is 78.0% (95% CI 58.6-100) in PD-L1 positive patients. Events: 3/15 (20%).", 90
  ), collapse = "\n")

pfs_pdl1$plot <- pfs_pdl1$plot + annotate(
  "text",
  x = 0.5, y = 0.08,
  label = caption2,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

pfs_pdl1$plot <- pfs_pdl1$plot + annotate(
  "text",
  x = 0.5, y = 0.18,
  label = caption3,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

### Save plot ---------------------------------------------------------------
ggsave(
  paste("../output/", project_id, "_5-mo_PFS_PD-L1_", as.Date(Sys.Date()), ".png", sep = ""),
  pfs_pdl1,
  width = 20,
  height = 10,
  units = "cm",
  dpi = 300
)

### Show plot --------------------------------------------------------------
pfs_pdl1
```

#### **Table 4. PFS by complete dose of Lenvatinib during the first 8 weeks:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
var_label(pfs$per_protocol_set) <- "Complete lenvatinib dose first 8 weeks"

tbl_survfit(
  survfit(Surv(pfs, pfs_event) ~ per_protocol_set, data = pfs, conf.type = "log-log"),
  times = c(5, 12),
  label_header = "**{time}-Month PFS**"
) %>%
  modify_header(label = " ")
```

```{r pfs-lenvatinib-dose-kaplan-meier-plot, message = FALSE, warning = FALSE, echo = FALSE}
### Plot ----------------------------------------------------------------------
pfs_pp <- ggsurvplot(
  fit = survfit(Surv(pfs, pfs_event) ~ per_protocol_set, data = pfs),
  xlab = "Months since treatment initiation",
  ylab = "Progression-Free Survival",
  conf.int = FALSE,
  pval = TRUE,
  pval.size = 3,
  pval.coord = c(0.5, 0.28),
  xlim = c(0, 24.9),
  ylim = c(0, 1.009),
  censor.size = 3.5,
  size = 1,
  break.time.by = 3,
  axes.offset = FALSE,
  surv.scale = "percent",
  break.y.by = 0.20,
  risk.table = TRUE,
  fontsize = 2.5,
  ggtheme = theme_classic(),
  palette = c("#BB322F", "#0A3158"),
  legend.labs = c("<8 weeks", "≥8 weeks"),
  legend.title = " "
)

### Modify theme --------------------------------------------------------------
pfs_pp <- customize_labels(
  pfs_pp,
  font.title     = c(8, "bold", "#0a0908"),
  font.caption   = c(8, "plain", "#0a0908"),
  font.x         = c(8, "bold", "#0a0908"),
  font.y         = c(8, "bold", "#0a0908"),
  font.xtickslab = c(8, "plain", "#0a0908"),
  font.ytickslab = c(8, "plain", "#0a0908")
)

grid.draw.ggsurvplot <- function(x) {
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

### Add legend ----------------------------------------------------------------
caption2 <- paste(strwrap(
  "5-month PFS is 71.3% (95% CI 44.0-87.0) with <8 weeks of complete Lenvatinib dose. Events: 5/19 (26.3%).", 120
  ), collapse = "\n")

caption3 <- paste(strwrap(
  "5-month PFS is 100% with ≥8 weeks of complete Lenvatinib dose. Events: 0/24 (0%).", 100
  ), collapse = "\n")

pfs_pp$plot <- pfs_pp$plot + annotate(
  "text",
  x = 0.5, y = 0.08,
  label = caption2,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

pfs_pp$plot <- pfs_pp$plot + annotate(
  "text",
  x = 0.5, y = 0.18,
  label = caption3,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

### Save plot ---------------------------------------------------------------
ggsave(
  paste("../output/", project_id, "_5-mo_PFS_PP_", as.Date(Sys.Date()), ".png", sep = ""),
  pfs_pp,
  width = 20,
  height = 10,
  units = "cm",
  dpi = 300
)

### Show plot --------------------------------------------------------------
pfs_pp
```

## Overall survival (OS)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
os <- read_excel(
  here(paste0("data/", "MedOPP341_PECATI_Draft_Listings_", cutoff_date, ".xlsx")),
  sheet = "LIST18"
) %>%
  clean_names() %>% 
  rename(
    patient = subject_id,
    os_event = overall_survival_os,
    os = time_of_os_months
  ) %>% 
  mutate(
    os_event = if_else(
      os_event == "No", 0, 1
    )
  )
```

#### **Table 5. OS:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
tbl_survfit(
  survfit(Surv(os, os_event) ~ 1, data = os, conf.type = "log-log"),
  times = c(5, 12),
  label_header = "**{time}-Month OS**"
) %>%
  modify_header(label = " ")
```

```{r os-kaplan-meier-plot, message = FALSE, warning = FALSE, echo = FALSE}
### Plot ----------------------------------------------------------------------
os_secondary <- ggsurvplot(
  fit = survfit(Surv(os, os_event) ~ 1, data = os),
  xlab = "Months since treatment initiation",
  ylab = "Overall Survival",
  conf.int = FALSE,
  xlim = c(0, 26.9),
  ylim = c(0, 1.009),
  censor.size = 3.5,
  size = 1,
  break.time.by = 3,
  axes.offset = FALSE,
  surv.scale = "percent",
  break.y.by = 0.20,
  risk.table = TRUE,
  fontsize = 2.5,
  ggtheme = theme_classic(),
  palette = c("#0A3158"),
  legend.labs = c(" "),
  legend.title = " "
)

### Modify theme --------------------------------------------------------------
os_secondary <- customize_labels(
  os_secondary,
  font.title     = c(8, "bold", "#0a0908"),
  font.caption   = c(8, "plain", "#0a0908"),
  font.x         = c(8, "bold", "#0a0908"),
  font.y         = c(8, "bold", "#0a0908"),
  font.xtickslab = c(8, "plain", "#0a0908"),
  font.ytickslab = c(8, "plain", "#0a0908")
)

grid.draw.ggsurvplot <- function(x) {
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

### Add legend ----------------------------------------------------------------
caption4 <- paste(strwrap(
  "5-month OS is 95.3% (95% CI 82.5-98.8). Events: 2/43 (4.7%).", 90
  ), collapse = "\n")

caption5 <- paste(strwrap(
  "12-month OS is 85.2% (95% CI 67.4-93.7). Events: 5/43 (11.6%).", 90
  ), collapse = "\n")

pfs_pdl1$plot <- pfs_pdl1$plot + annotate(
  "text",
  x = 0.5, y = 0.08,
  label = caption2,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

pfs_pdl1$plot <- pfs_pdl1$plot + annotate(
  "text",
  x = 0.5, y = 0.18,
  label = caption3,
  cex = 3, vjust = "center", hjust = "left", fontface = 20
)

### Save plot ---------------------------------------------------------------
ggsave(
  paste("../output/", project_id, "_12-mo_OS_global_", as.Date(Sys.Date()), ".png", sep = ""),
  os_secondary,
  width = 20,
  height = 10,
  units = "cm",
  dpi = 300
)

### Show plot --------------------------------------------------------------
os_secondary
```

## Best Overall Response

```{r, message = FALSE, warning = FALSE, echo = FALSE}
### Load data -----------------------------------------------------------------
data_waterfall <- read_excel(
  here(paste0("data/", "MedOPP341_PECATI_Draft_Listings_", cutoff_date, ".xlsx")),
  sheet = "LIST18"
  ) %>% 
  clean_names() %>% 
  rename(
    patient = subject_id,
    confirmed_bor = confirmed_best_overall_response_bor,
    sld_cfb = maximum_tumor_shrinkage_percent
  ) %>% 
  mutate(
    sld_cfb = if_else(
      is.na(sld_cfb), 1, sld_cfb
    ),
    confirmed_bor = if_else(
      confirmed_bor == "SD>=24w", "SD", confirmed_bor
    ),
    confirmed_bor = if_else(
      confirmed_bor == "SD<24w", "SD", confirmed_bor
    ),
    sld_cfb = if_else(
      sld_cfb == 0, 1, sld_cfb
    )
  ) %>%
  arrange(-sld_cfb) %>%
  mutate(
    patient = factor(patient),
    confirmed_bor = factor(confirmed_bor, levels = c("CR", "PR", "SD", "PD", "NE")),
    order = row_number(),
    colour = if_else(
      confirmed_bor == "CR", "#111111",
      if_else(
        confirmed_bor == "PR", "#152546",
          if_else(
            confirmed_bor == "SD", "#666F1C",
            if_else(
              confirmed_bor == "PD", "#99001B",
              "#717272"
            )
          )
        )
      )
    ) %>%
  select(
    patient, group, pd_l1, per_protocol_set, confirmed_bor, sld_cfb, colour, order
  )

### Plot ----------------------------------------------------------------------
waterfallplot_orr_confirmed <- ggplot(
  data_waterfall,
  aes(
    x = order,
    y = sld_cfb,
    colour = factor(colour)
    )
  ) +
  coord_cartesian(
    ylim = c(-100, 50)
    ) +
  scale_y_continuous(
    breaks = seq(-100, 50, by = 20
                 )
    ) +
  geom_bar(
    fill = data_waterfall$colour,
    colour = "#FFFFFF",
    stat = "identity",
    width = 0.99
    ) +
  geom_abline(slope = 0, intercept = 20, col = "gray40", lty = 2) +
  geom_abline(slope = 0, intercept = -30, col = "gray40", lty = 2) +
  theme_classic() +
  labs(title = "",
       x = "Patients",
       y = "Maximum change from baseline in tumor size") +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    )

### Save plot -----------------------------------------------------------------
ggsave(
  here("output/PECATI_ORR_confirmed_waterfall_", cutoff_date, ".png"),
  waterfallplot_orr_confirmed,
  width = 16,
  height = 8,
  units = "cm",
  dpi = 300
  )

### Show plot -----------------------------------------------------------------
waterfallplot_orr_confirmed
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_12_or_recist <- read_excel(
  here(paste0("output/", project_id, "_12_or_recist_2024-06-20", ".xlsx"))
  )

pdl1 <- read_excel(
  here(paste0("data/", "PECATI_PDL1_", cutoff_date, ".xlsx")),
  col_types = c("text", rep("skip", 6), "text")
  )

pdl1 <- pdl1 %>% 
  clean_names() %>% 
  mutate(
    pdl1_percent = if_else(is.na(pdl1_percent), "Unknown", pdl1_percent),
    patient = paste0("0", patient)
  )

data_12_or_recist <- merge(
  pdl1,
  data_12_or_recist,
  by = "patient",
  all = TRUE
  )

data_12_or_recist <- data_12_or_recist %>%
  mutate(
    pdl1_percent = if_else(is.na(pdl1_percent), "Unknown", pdl1_percent)
    # pdl1_percent = factor(pdl1_percent, levels = c("Unknown", "0", "1 <- 49", "≥50")),
  )

data_12_or_recist <- data_12_or_recist %>%
  mutate(
    event_num = as.numeric(event_num),
    percent_change_from_baseline = as.numeric(percent_change_from_baseline),
    sd_this_ta = as.numeric(sd_this_ta),
    pr_this_ta = as.numeric(pr_this_ta),
    cr_this_ta = as.numeric(cr_this_ta),
    pd_this_ta = as.numeric(pd_this_ta)
  )


spider_plot <- function() {
  spider_ta <- data_12_or_recist %>%
    ggplot(
      aes(x = event_num, y = percent_change_from_baseline, group = patient)
    ) +
    # Line plot showing SDL percent change from baseline, colored by pdl1_percent and grouped by patient
    geom_line(color = "#2C2C2B", size = 0.75) +  # aes(color = pdl1_percent), 
    # Second data layer: dot plot showing TAs with an OR = SD
    geom_point(
      aes(
        x = sd_this_ta
      ),
      size = 1.5,
      stroke = 1.5,
      shape = 16
    ) +
    # Third data layer: dot plot showing TAs with an OR = PR
    geom_point(
      aes(
        x = pr_this_ta
      ),
      size = 1.5,
      stroke = 1.5,
      shape = 16
    ) +
    # Fourth data layer: dot plot showing TAs with an OR = CR
    geom_point(
      aes(
        x = cr_this_ta
      ),
      size = 1.5,
      stroke = 1.5,
      shape = 16
    ) +
    # Fifth data layer: dot plot showing TAs with an OR = PD
    geom_point(
      aes(
        x = pd_this_ta,
      ),
      color = "#E63C56",
      size = 1.5,
      stroke = 1.5,
      shape = 3
    ) +
    # Final aesthetics adjustments
    theme_minimal() +
    theme(
      panel.border = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.background = element_rect(fill = "#FFFFFF", colour = "#FFFFFF"),
      plot.background = element_rect(fill = "#FFFFFF", colour = "#FFFFFF"),
      legend.title = element_text(size = 5),
      legend.text = element_text(size = 5),
      legend.justification = "top",
      legend.position = c(0.95, 0.95),
      legend.direction = "vertical"
    ) +
    labs(
      x = "Tumor assessment",
      y = "Change from baseline in tumor size (%)",
      shape = "Overall response"
    ) +
    guides(
      size = "none",
      color = guide_legend(title = "PD-L1 (%)")
    ) +
    scale_x_continuous(breaks = seq(1, 11, 1)) +
    scale_y_continuous(breaks = seq(-100, 100, 20))

  spider_ta <- spider_ta + scale_shape(na.translate = FALSE)

  ggsave(
    paste("../output/", project_id, "_spider_TAs_", as.Date(Sys.Date()), ".png", sep = ""),
    spider_ta,
    width = 20,
    height = 10,
    units = "cm",
    dpi = 300
  )

  return(spider_ta)
}

spider_plot()
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Number of patients with PFS event at 5 months
num_events_5_months <- 5

# Total number of patients
total_patients <- 43

# Observed 5-month PFS rate
observed_pfs_rate <- (total_patients - num_events_5_months) / total_patients

# Hypothesized 5-month PFS rate under null hypothesis H0
hypothesized_pfs_rate <- 0.50

# Perform a one-sided binomial test
binom_test <- binom.test(total_patients - num_events_5_months, total_patients, p = hypothesized_pfs_rate, alternative = "greater")

# Extract p-value
p_value <- binom_test$p.value

# Check if the p-value is less than 0.05
if (p_value < 0.05) {
  cat("The p-value is less than 0.05\n")
  cat("We have a statistically significant result\n")
} else {
  cat("The p-value is not less than 0.05\n")
  cat("We do not have a statistically significant result\n")
}

# Print the observed 5-month PFS rate and p-value
cat("Observed 5-month PFS rate:", observed_pfs_rate, "\n")
cat("p-value:", p_value, "\n")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
data_temp_reduction <- read_excel(
  here(paste0("data/", "MedOPP341_PECATI_Draft_Listings_", cutoff_date, ".xlsx")),
  sheet = "LIST19"
  ) %>% 
  clean_names() %>% 
  rename(
    patient = subject_id,
    pp = per_protocol_set,
    reduction = study_treatment_dose_reductions
  ) %>% 
  select(
  patient, pp, reduction
  )

count_table <- table(data_temp_reduction$pp, data_temp_reduction$reduction)
percentage_table <- prop.table(count_table) * 100
combined_table <- paste(count_table, "(", round(percentage_table, 1), "%)", sep="")

combined_table_matrix <- matrix(combined_table, nrow = nrow(count_table), ncol = ncol(count_table))
rownames(combined_table_matrix) <- paste("Lenvatinib 20mg ≥8 weeks:", rownames(count_table))
colnames(combined_table_matrix) <- paste("Lenvatinib dose reduction due to toxicity:", colnames(count_table))

combined_table_matrix
```

#### **Table 6. PFS inm patients with PD-L1 ≥50%:**

```{r, message = FALSE, warning = FALSE, echo = FALSE}
pfs_pdl1_50 <- merge(
  pdl1,
  pfs,
  by = "patient",
  all = TRUE
  ) %>% 
  filter(
    pdl1_percent == "≥50"
  )

tbl_survfit(
  survfit(Surv(pfs, pfs_event) ~ 1, data = pfs_pdl1_50, conf.type = "log-log"),
  times = c(5, 12),
  label_header = "**{time}-Month PFS**"
) %>%
  modify_header(label = " ")
```
